### 2024-03-11 paper/paper.qmd
Initial commit

---
title: "Operating Budget Program Summary By Expensiture Category"
subtitle: "Murrumbidgee Paper"
execute:
  echo: false
  message: false
  warning: false
author: 
  - Joe Zhou
thanks: "Code and data are available at: LINK."
date: today
date-format: long
mainfont: "Hiragino Sans GB"
format:
  pdf:
    documentclass: scrartcl
    geometry:
      - top=30mm
      - left=20mm
      - heightrounded
    fontfamily: libertinus
    colorlinks: true
    toc: true
classoption: abstract
abstract: | 
  text
include_in_header:
  text:
    \renewcommand{\abstractname}{Abstract}
bibliography: references.bib
---

```{r}
library(dplyr)
library(modelsummary)
library(tidyverse)
library(ggplot2)
library(rstanarm)
library(here)
library(arrow)
library(tidybayes)
library(tidyverse)
library(marginaleffects)
library(readxl)
library(httr)
```

```{r}
#| output: false

```

# Introduction

# Data

# Model

# Results

# Discussion

\newpage

# References
@citeR

### 2024-03-11 scripts/00-simulate_data.R

@@ -1,0 +1,196 @@
+#### Workspace setup ####
+library(dplyr)
+library(modelsummary)
+library(tidyverse)
+library(ggplot2)
+library(rstanarm)
+library(here)
+library(arrow)
+library(tidybayes)
+library(tidyverse)
+library(marginaleffects)
+#### Simulate data ####
+input_file_path <- here("data", "analysis_data","cleaned_data.parquet")
+input_file_path_data <- here("data", "analysis_data","data.parquet")
+
+cleaned_data <- read_parquet(input_file_path)
+data <- read_parquet(input_file_path_data)
+
+
+
+cleaned_data |> 
+  ggplot(aes(x=Program, y=Amount, alpha=0.001)) +
+  geom_point() +
+  geom_abline(intercept = 0, slope = 5, linetype = "dashed") +
+  theme_classic() +
+  scale_x_log10() + 
+  scale_y_log10()
+
+cleaned_data |> 
+  ggplot(aes(x=Service, y=Amount, alpha=0.11)) +
+  geom_point() +
+  scale_x_log10() + 
+  scale_y_log10()
+
+cleaned_data |> 
+  ggplot(aes(x=CategoryName, y=Amount, alpha=0.11)) +
+  geom_point() +
+  scale_x_log10() + 
+  scale_y_log10()
+
+
+
+
+
+
+
+
+
+
+
+base_plot <- 
+  cleaned_data |>
+  ggplot(aes(x = Program, y = Amount)) +
+  geom_point(alpha = 0.5) +
+  labs(
+    x = "Service",
+    y = "Amount"
+  ) +
+  theme_classic()
+
+base_plot +
+  geom_smooth(
+    method = "lm",
+    se = TRUE,
+    color = "black",
+    linetype = "dashed",
+    formula = "y ~ x"
+  )
+
+
+
+
+
+
+
+reg_amount_program <-
+  stan_glm(
+    formula = Amount ~ Program,
+    data = cleaned_data,
+    family = gaussian,
+    prior = normal(location = 8, scale = 2.5, autoscale = TRUE),
+    prior_intercept = normal(0, 2.5, autoscale = TRUE),
+    prior_aux = exponential(rate = 1, autoscale = TRUE),
+    seed = 853
+  )
+  # saveRDS(
+  #   sim_run_data_second_model_rstanarm,
+  #   file = "sim_run_data_second_model_rstanarm.rds"
+  # )
+
+
+  
+
+reg_amount_program_expenseOrRevenue <-
[…]

### 2024-03-11 scripts/01-download_data.R

@@ -1,0 +1,22 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from OpenDataToronto
+# Author: ZeWei Zhou
+# Date: March 8, 2024
+# Contact: zewei.zhou@mail.utoronto.ca
+# License: MIT
+
+#### Workspace setup ####
+library(httr)
+library(here)
+
+#### Download data ####
+url = "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/2c90a5d3-5598-4c02-abf2-169456c8f1f1/resource/d55a2458-f116-456e-a3be-4a0d867fa190/download/approved-operating-budget-summary-2016.xlsx"
+
+#### Save data ####
+# Specify the path where the downloaded file will be saved
+save_to <- here("data", "raw_data", "raw_data.xlsx")
+
+# Download the file by sending an HTTP request GET to the url and write to the 
+# disk with the specified path
+GET(url, write_disk(save_to, overwrite = TRUE))
+

### 2024-03-11 scripts/02-data_cleaning.R

@@ -1,0 +1,134 @@
+#### Preamble ####
+# Purpose: Clean the downloaded raw data
+# Author: ZeWei Zhou
+# Date: March 8, 2024
+# Contact: zewei.zhou@mail.utoronto.ca
+# License: MIT
+
+#### Workspace setup ####
+library(dplyr)
+library(readxl)
+library(here)
+library(arrow)
+
+#### Clean data ####
+input_file_path <- here("data", "raw_data", "raw_data.xlsx")
+output_file_path_analysis_parquet <- here("data", "analysis_data", "cleaned_data.parquet")
+output_file_path_parquet <- here("data", "analysis_data", "data.parquet")
+output_file_path_csv <- here("data", "analysis_data", "cleaned_data.csv")
+
+raw_data <- readxl::read_xlsx(input_file_path)
+cleaned_data = raw_data
+
+colnames(cleaned_data)[3] <- "CategoryName"
+colnames(cleaned_data)[4] <- "ExpenseOrRevenue"
+colnames(cleaned_data)[5] <- "Amount"
+
+Program_list <- unique(cleaned_data$Program)
+Service_list <- unique(cleaned_data$Service)
+CategoryName_list <- unique(cleaned_data$CategoryName)
+ExpenseOrRevenue_list <- unique(cleaned_data$ExpenseOrRevenue)
+
+
+
+for(i in 1:length(cleaned_data$Program)){
+  cleaned_data$Program[i] = grep(cleaned_data$Program[i], Program_list)
+}
+for(i in 1:length(cleaned_data$Service)){
+  cleaned_data$Service[i] = grep(cleaned_data$Service[i], Service_list)
+}
+for(i in 1:length(cleaned_data$CategoryName)){
+  cleaned_data$CategoryName[i] = grep(cleaned_data$CategoryName[i], CategoryName_list)
+}
+for(i in 1:length(cleaned_data$ExpenseOrRevenue)){
+  cleaned_data$ExpenseOrRevenue[i] = grep(cleaned_data$ExpenseOrRevenue[i], ExpenseOrRevenue_list)
+}
+
+
+
+
+
+# cleaned_data <- 
+#   raw_data |>
+#   mutate(
+#     Program = case_when(
+#       Program == '311 Toronto' ~ 1,
+#       Program == 'Affordable Housing Office' ~ 2,
+#       Program == "Arena Boards of Management" ~ 3,
+#       Program == "Association of Community Centres" ~ 4,
[…]

### 2024-03-12 paper/datasheet_template.qmd
Init

---
title: "Datasheet for 'Operating Budget Program Summary By Expenditure Category'"
subtitle: "Murrumbidgee Paper"
author: 
  - ZeWei Zhou
thanks: |
  Code and data supporting this analysis is avaliable at:https://github.com/
  zhouJoe2033/MurrumbidgeePaper/blob/main"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to enable analysis of Australian politicians. We were unable to find a publicly available dataset in a structured format that had the biographical and political information on Australian politicians that was needed for modelling.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Open Data Toronto
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - Open Data Toronto
4. *Any other comments?*
    - I am not sure yet

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- documents, yes, ok
2. *How many instances are there in total (of each type, if appropriate)?*
	- a lot, from year 2024 to year 2018
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- TBD
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- TBD
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- TBD
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- TBD
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- TBD
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- TBD
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- TBD
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- TBD
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- TBD
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- TBD
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- TBD
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- TBD
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- TBD
16. *Any other comments?*
	- TBD

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

\newpage

# References

### 2024-03-12 paper/paper.qmd
Init

@@
 ---
-title: "Operating Budget Program Summary By Expensiture Category"
+title: "Operating Budget Program Summary By Expenditure Category"
 subtitle: "Murrumbidgee Paper"
@@
   - Joe Zhou
-thanks: "Code and data are available at: LINK."
+thanks: |
+  Code and data supporting this analysis is avaliable at:https://github.com/
+  zhouJoe2033/MurrumbidgeePaper/blob/main"
 date: today
@@
 abstract: | 
-  text
+  This analysis paper downloaded data from the OpenDataToronto website, data is 
+  randomly choosen at the year of 2016 to proceed analyasis. The dataset 
+  illustrates the approoved and recommened annual operating budget summary by 
+  expenditue category in each program. This paper found that program is 
+  positively correlated to the budget amount, and expense or revenue and 
+  category name is negatively correlated to the buget amount. This found matters
+  because it shows that the buget amount was set to a high level at first, and
+  then if the program adds value to the buget amount depends on which program 
+  is applied, and weather the the application is expense or revenue is likely 
+  devaluing the approved budget amount by a large scale, and the category name 
+  is likely devaluing the approved budget amount by a small scale.However, since
+  the R squared value is not significant and the adjusted R squared value is 
+  even less significant, this dataset requires more variables to determine the
+  accurate relationship among variables and the budget amount.
+
 include_in_header:
@@
 ```{r}
+library(arrow)
 library(dplyr)
@@
 library(here)
-library(arrow)
 library(tidybayes)
@@
 #| output: false
+#### File path ####
+input_file_path_cleaned_data <- 
+  here("data", "analysis_data","cleaned_data.parquet")
 
+reg_amount_program_path <- 
+  here("models", "reg_amount_program_expenseOrRevenue.rds")
+reg_amount_program_expenseOrRevenue_path <- 
+  here("models", "reg_amount_program_expenseOrRevenue.rds")
+reg_amount_program_expenseOrRevenuw_categoryName_path <- 
+  here("models", "reg_amount_program_expenseOrRevenuw_categoryName.rds")
+reg_amount_program_expenseOrRevenuw_categoryName_service_path <- 
+  here("models", "reg_amount_program_expenseOrRevenuw_categoryName_service.rds")
+
+#### Retrieve Data ####
+cleaned_data <- read_parquet(input_file_path_cleaned_data)
+
+#### Retrieve Model ####
+reg_amount_program = readRDS(reg_amount_program_path)
+reg_amount_program_expenseOrRevenue = 
+  readRDS(reg_amount_program_expenseOrRevenue_path)
+reg_amount_program_expenseOrRevenuw_categoryName = 
+  readRDS(reg_amount_program_expenseOrRevenuw_categoryName_path)
+reg_amount_program_expenseOrRevenuw_categoryName_service = 
+  readRDS(reg_amount_program_expenseOrRevenuw_categoryName_service_path)
 ```
@@
 # Introduction
+Welcome to the City of Toronto's Open Data platform, where transparency meets information accessibility.
+In our ongoing commitment to fostering an informed and engaged community, we present the "Budget Operating Budget Program Summary by Expenditure Category" dataset.
+This comprehensive resource provides a detailed overview of the city's financial allocations, offering a transparent glimpse into the intricacies of budget distribution across various expenditure categories.
+This analysis paper collects data from the City of Toronto's Open Data and uses several statistic techniques in order to process professional analysis of how provided variables in the dataset is correlated to the budget amount, and using the statistic model in order to make predictions of budget amount based on different approached and models.
+And illustrates these relationship by using graphs and tables to the readers.
+
+Provided dataset in the year of 2016 has five variables in total and 1313 observations, however, the amount of observations is not good enough to make predictions, and the statistic results shows that there exists more variables that are correlated to the budget amount variable.
+This paper uses several approaches to build different regression models in order to analysis different statistic results, and using these statistic results to build graphs and tables to illustrates to the readers to understand the relationship among the budget amount and the rest of the variables.
+
+This paper found that program is positively correlated to the budget amount, and expense or revenue and category name is negatively correlated to the buget amount.
+This found matters because it shows that the budget amount was set to a high level at first, and then if the program adds value to the buget amount depends on which program is applied, and weather the the application is expense or revenue is likely devaluing the approved budget amount by a large scale, and the category name is likely devaluing the approved budget amount by a small scale.
+The statistic results by using different approaches to process is important since the results tells the R squared value is not significant and the adjusted R squared value is even less significant, and thus the conclusion is this dataset requires more variables to determine the accurate relationship among variables and the budget amount.
+
+After the introduction of this paper, the data section will be well explained in the next section and the model will be illustrated as following, then the statistic tables will be showing and the discussion section will be written to make conclusions of the founding of this paper.
+At last, delve into the heart of Toronto's fiscal landscape, exploring the nuanced breakdown of operating budgets for different programs.
+This dataset serves as a valuable tool for researchers, policymakers, and citizens alike, empowering them with the knowledge needed to understand how public funds are allocated to support essential services and programs.
+As we navigate the complexities of municipal governance, this dataset acts as a beacon of openness, promoting accountability and understanding.
+The information contained within invites you to explore the financial decisions that shape the city's growth and well-being.
+Join us on this data-driven journey towards a more informed and engaged Toronto.
+
+```{r}
+#| label: fig-1-123
+#| fig-cap: "Data Plot of Program, Category Name and Expense Or Revenue to the Amount"
+#| fig-subcap: ["Data Plot of Program to the Amount","Data Plot of Program to the Category","Data Plot of Program to the Expense Or Revenue"]
+#| layout-ncol: 2
+#| layout-nrow: 2
+cleaned_data |> 
+  ggplot(aes(x=Program, y=Amount, alpha=0.1)) +
+  geom_point() +
+  geom_abline(intercept = 0, slope = 5, linetype = "dashed") +
+  theme_classic() +
+  scale_x_log10() + 
+  scale_y_log10()
+
+cleaned_data |> 
+  ggplot(aes(x=CategoryName, y=Amount, alpha=0.1)) +
+  geom_point() +
+  theme_classic() +
+  scale_x_log10() + 
+  scale_y_log10()
+
+cleaned_data |> 
+  ggplot(aes(x=ExpenseOrRevenue, y=Amount, alpha=0.1)) +
+  geom_point() +
+  theme_classic() +
+  scale_x_log10() + 
+  scale_y_log10()
+
+base_plot <- 
+  cleaned_data |>
+  ggplot(aes(x = Program, y = Amount)) +
+  geom_point(alpha = 0.5) +
+  labs(
+    x = "Service",
+    y = "Amount"
+  ) +
+  theme_classic()
+
+base_plot +
+  geom_smooth(
+    method = "glm",
+    se = TRUE,
+    color = "black",
+    linetype = "dashed",
+    formula = "y ~ x"
+  )
+```
 
 # Data
+The "Approved Operating Budget Summary 2016" dataset offers a detailed and structured insight into the financial framework that governed the City of Toronto during the fiscal year 2016.
+Presented in an Excel format, this dataset encapsulates a comprehensive overview of the city's approved operating budget, providing a granular breakdown across various categories and programs.
+
+Contained within this dataset are meticulously organized details pertaining to revenue sources, expenditure allocations, and program-specific budgetary considerations.
+Researchers, policymakers, and citizens alike can leverage this information to gain a deeper understanding of how financial resources were strategically distributed to support essential services, initiatives, and projects throughout the city during the specified fiscal period.
+
+Whether you are interested in analyzing trends, assessing the prioritization of different programs, or simply seeking a transparent view of municipal financial decision-making, this dataset serves as a valuable resource.
+It contributes to the broader goals of transparency, accountability, and citizen engagement, fostering a more informed and empowered community.
+Explore the "Approved Operating Budget Summary 2016" dataset to uncover the financial intricacies that shaped the city's landscape during that specific fiscal year, and gain valuable insights into the priorities and allocations that drove Toronto's development and well-being.
+
+This cleaned dataset has three variables that is correlated to the budget amount variable such that @fig-1-123-1 illustrates the program type through a data plot that shows the 53 programs in total and the amount of each program is budged, a lot of the program is over 1e+03 and most of the approved program is ranked after 30(index number based on the program type).@fig-1-123-2 illustrates the category type through a data plot that shows the 17 programs in total and the amount of each category is budged, the amount that is approved by the category name is vertically and almost behaves like a line, that means there are categories is not approved and the approved category is correlated by other variables therefore it causes they have different approved budget amount.@fig-1-123-3 illustrates the budget or revenue type through a data plot that shows the 2 types in total and the amount of each type is budged, apparently, most of the budget approved is revenue and only a few of them is expense.@fig-1-123-4 shows the Service variable and its relationship with the Amount bariable, and most of it shows at zero which indicates a error of the data.
+
+Through the open data website, there are other years of operating budget that can be found, each year the dataset gets improved by adding more variables and more observations.
+It takes time to collect a well organized dataset and it takes time to make a dataset useful and meaningful.
+The data barely has null value therefore it is a good news to the high-level cleaning since converting from text to numeric and analysis the data in statistic model is good enough, however, scripts wrote in R functions well does not mean it functions well in the .qmd file, and sometimes the code gets messy and does not work properly when the compiler shows error.
+
+Overall, the observation number is 1312 and the amount variable has a minimum number -3.931e+09 and a maximum number 1.273e+09, the median of the amount variable is 3.583e+04 and the first quartile is -1.935e_05, and the third quartile of amount is 1.057e+06.
+The correlation relationship among these variables are Amount is positively correlated to the Program variable and Amount is negatively correlated to the ExpenseOrRevenue and Categoram Name variable.
 
 # Model
+The first model used in this paper is 
+$$
+\begin{aligned}
+Amount = beta_0 + beta_1*Program + epsilon
+\end{aligned}
+$$
+that  establish the relationship between Amount and Program variables, and it shows that beta_0 has a positive value at a large scale and the Program variable has a positive coefficient beta_1 that varies from the specific Program.
+The second model used in this paper is
+$$
+\begin{aligned}
+Amount = beta_0 + beta_1*Program + beta_2*ExpenseOrRevenue + epsilon
+\end{aligned}
+$$
+that establish the relationship between Amount,Program and ExpenseOrRevenue variables, and it shows that beta_0 has a positive value at a large scale and the Program variable has a positive coefficient beta_1 that varies from the specific Program and ExpenseOrRevenue has a negative coefficient beta_2 at a large scale.
+The third model used in this paper is
+$$
+\begin{aligned}
+Amount = beta_0 + beta_1*Program +  + beta_2*ExpenseOrRevenue + beta_3*CategoryName + epsilon
+\end{aligned}
+$$
+that establish the relationship between Amount,Program, ExpenseOrRevenue and CategoramName variables, and it shows that beta_0 has a positive value at a large scale and the Program variable has a positive coefficient beta_1 that varies from the specific Program and ExpenseOrRevenue has a negative coefficient beta_2 at a large scale and CategoryName variable has a negative coefficient beta_3 at a small scale.
+
+```{r}
+#| label: fig-2
+#| fig-cap: "Statistic Summary"
+```
 
@@
 
+Without direct access to the dataset, I can't provide specific statistical results.
+However, I can discuss potential statistical analyses and insights that could be derived from the "Approved Operating Budget Summary 2016" dataset:
+
+Descriptive Statistics: Descriptive statistics can provide an overview of the budgetary data.
+This includes measures such as the total budget, mean budget per program, standard deviation to assess variability, and range to understand the spread of budget allocations across different categories.
+Budget Allocation Trends: Analyzing budget allocations over time can reveal trends and patterns.
+You could calculate year-on-year changes in budget allocations for various programs or expenditure categories, identifying areas of growth or decline.
+Comparison of Budget Allocations: Statistical tests, such as ANOVA or t-tests, could be used to compare budget allocations across different departments, programs, or geographic regions within Toronto.
+This analysis could identify significant differences in funding priorities.
+Correlation Analysis: Explore correlations between budget allocations and other variables, such as population demographics, economic indicators, or service utilization rates.
+Understanding these correlations can provide insights into the factors influencing budget decisions.
+Forecasting: Time series analysis techniques could be employed to forecast future budget allocations based on historical data.
+This could help policymakers plan for future resource needs and identify potential budgetary challenges.
+Efficiency Analysis: Efficiency analysis methods, such as data envelopment analysis (DEA), could assess the efficiency of budget allocations in delivering desired outcomes or services.
+This analysis could highlight areas for improvement or reallocation of resources.
+Visualization: Visualizing budget data through charts, graphs, and maps can help communicate insights effectively.
+For example, bar charts can illustrate budget allocations for different programs, while maps can show spatial variations in funding across Toronto.
+By conducting these statistical analyses, stakeholders can gain valuable insights into the allocation, utilization, and impact of the operating budget in Toronto for the year 2016.
+These insights can inform decision-making, improve resource allocation efficiency, and ultimately contribute to better governance and service delivery.
+
 # Discussion
+The "Budget Operating Budget Program Summary by Expenditure Category" dataset hosted on the City of Toronto's Open Data platform provides a comprehensive and detailed breakdown of the city's operating budget across various expenditure categories.
+This dataset offers invaluable insights into how public funds are allocated to support different programs and services essential for the functioning and development of the city.
+
+Key features of this dataset include: Granular Expenditure Categories: The dataset categorizes expenditures into specific areas such as personnel, materials and supplies, services, transfers to reserves, and more.
+This granularity allows for a detailed analysis of where financial resources are directed within the city's operations.
+Program-Level Breakdown: Expenditures are further disaggregated by program, providing a clear understanding of how funds are distributed across different city initiatives.
+This level of detail enables stakeholders to assess the relative prioritization of various programs within the budget.
+Annual Budget Comparisons: The dataset may include information from multiple fiscal years, allowing for year-over-year comparisons and trend analysis.
+This feature is invaluable for identifying budgetary shifts, emerging priorities, and areas of consistency or change over time.
+
+Transparency and Accountability: By making this data publicly accessible, the City of Toronto promotes transparency in its financial decision-making processes.
+Citizens, researchers, and policymakers can scrutinize the budget allocations, fostering accountability and informed public discourse.
+Data Accessibility: The dataset is hosted on an open data platform, making it easily accessible to anyone with an internet connection.
+This accessibility democratizes access to crucial financial information, empowering individuals and organizations to engage with city finances in meaningful ways.
+
+Overall, the "Budget Operating Budget Program Summary by Expenditure Category" dataset serves as a foundational resource for understanding the financial landscape of the City of Toronto.
+It facilitates evidence-based decision-making, fosters transparency and accountability in governance, and empowers citizens to participate in shaping the future of their city.
 
@@
 @citeR
+@ChatGPT

### 2024-03-12 scripts/00-simulate_data.R

@@ -9,15 +9,35 @@ library(arrow)
 library(tidybayes)
 library(tidyverse)
 library(marginaleffects)
-#### Simulate data ####
-input_file_path <- here("data", "analysis_data","cleaned_data.parquet")
+
+
+#### Data Path ####
 input_file_path_data <- here("data", "analysis_data","data.parquet")
+input_file_path_cleaned_data <- here("data", "analysis_data","cleaned_data.parquet")
 
-cleaned_data <- read_parquet(input_file_path)
-data <- read_parquet(input_file_path_data)
+reg_amount_program_path <- 
+  here("models", "reg_amount_program_expenseOrRevenue.rds")
+reg_amount_program_expenseOrRevenue_path <- 
+  here("models", "reg_amount_program_expenseOrRevenue.rds")
+reg_amount_program_expenseOrRevenuw_categoryName_path <- 
+  here("models", "reg_amount_program_expenseOrRevenuw_categoryName.rds")
+reg_amount_program_expenseOrRevenuw_categoryName_service_path <- 
+  here("models", "reg_amount_program_expenseOrRevenuw_categoryName_service.rds")
 
 
+#### Read data ####
+cleaned_data <- read_parquet(input_file_path_cleaned_data)
 
+reg_amount_program = readRDS(reg_amount_program_path)
+reg_amount_program_expenseOrRevenue = 
+  readRDS(reg_amount_program_expenseOrRevenue_path)
+reg_amount_program_expenseOrRevenuw_categoryName = 
+  readRDS(reg_amount_program_expenseOrRevenuw_categoryName_path)
+reg_amount_program_expenseOrRevenuw_categoryName_service = 
+  readRDS(reg_amount_program_expenseOrRevenuw_categoryName_service_path)
+
+
+#### Simulate data ####
 cleaned_data |> 
   ggplot(aes(x=Program, y=Amount, alpha=0.001)) +
   geom_point() +
@@ -27,26 +47,16 @@ cleaned_data |>
   scale_y_log10()
 
 cleaned_data |> 
-  ggplot(aes(x=Service, y=Amount, alpha=0.11)) +
-  geom_point() +
-  scale_x_log10() + 
-  scale_y_log10()
-
-cleaned_data |> 
   ggplot(aes(x=CategoryName, y=Amount, alpha=0.11)) +
   geom_point() +
   scale_x_log10() + 
   scale_y_log10()
 
-
-
-
-
-
-
-
-
-
[…]

### 2024-03-12 scripts/02-data_cleaning.R

@@ -29,23 +29,26 @@ Service_list <- unique(cleaned_data$Service)
 CategoryName_list <- unique(cleaned_data$CategoryName)
 ExpenseOrRevenue_list <- unique(cleaned_data$ExpenseOrRevenue)
 
-
-
 for(i in 1:length(cleaned_data$Program)){
   cleaned_data$Program[i] = grep(cleaned_data$Program[i], Program_list)
 }
 for(i in 1:length(cleaned_data$Service)){
-  cleaned_data$Service[i] = grep(cleaned_data$Service[i], Service_list)
+    cleaned_data$Service[i] = grep(cleaned_data$Service[i], Service_list)
 }
 for(i in 1:length(cleaned_data$CategoryName)){
-  cleaned_data$CategoryName[i] = grep(cleaned_data$CategoryName[i], CategoryName_list)
+  cleaned_data$CategoryName[i] = grep(cleaned_data$CategoryName[i], 
+                                      CategoryName_list)
 }
 for(i in 1:length(cleaned_data$ExpenseOrRevenue)){
-  cleaned_data$ExpenseOrRevenue[i] = grep(cleaned_data$ExpenseOrRevenue[i], ExpenseOrRevenue_list)
+  cleaned_data$ExpenseOrRevenue[i] = grep(cleaned_data$ExpenseOrRevenue[i], 
+                                          ExpenseOrRevenue_list)
 }
 
 
-
+cleaned_data$Program <- as.numeric(cleaned_data$Program)
+cleaned_data$CategoryName <- as.numeric(cleaned_data$CategoryName)
+cleaned_data$Service <- as.numeric(cleaned_data$Service)
+cleaned_data$ExpenseOrRevenue <- as.numeric(cleaned_data$ExpenseOrRevenue)
 
 
 # cleaned_data <- 
@@ -114,12 +117,6 @@ for(i in 1:length(cleaned_data$ExpenseOrRevenue)){
 #   ) |>
 #   select(Program, Service)
 
-
-
-
-
-
-  
 
 
 

### 2024-03-12 scripts/03-test_data.R

@@ -1,15 +1,51 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests
+# Author: ZeWei Zhou
+# Date: 12 March 2024
+# Contact: zewei.zhou@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(testthat)
+#### Data Path ####
+input_file_path_data <- here("data", "analysis_data","data.parquet")
+
+#### Read data ####
+cleaned_data <- read_parquet(input_file_path_cleaned_data)
 
 #### Test data ####
+test_that("Check class", { expect_type(cleaned_data$Program, "double") })
+test_that("Check class", { expect_type(cleaned_data$Service, "double") })
+test_that("Check class", { expect_type(cleaned_data$CategoryName, "double") })
+test_that("Check class", { expect_type(cleaned_data$ExpenseOrRevenue,"double")})
+test_that("Check class", { expect_type(cleaned_data$Amount,"double") })
+
+test_that("Check number of observations is correct", { 
+  expect_equal(nrow(cleaned_data), 1312)
+})
+test_that("Check complete", {
+  expect_true(all(complete.cases(cleaned_data)))
+})
+
+
+test_that("Check coefficients", { 
+  expect_gt(cleaned_data$Program[0], 0) })
+test_that("Check coefficients", { 
+  expect_gt(cleaned_data$Service[1], 10) })
+test_that("Check coefficients", { 
+  expect_gt(cleaned_data$CategoryName[2], 20) })
+test_that("Check coefficients", { 
+  expect_gt(cleaned_data$ExpenseOrRevenue[3], 30) })
+test_that("Check coefficients", { 
+  expect_gt(cleaned_data$Amount[4], 30) })
+
+
+
+
+
+
+
+
+

### 2024-03-12 scripts/04-model.R

@@ -10,28 +10,91 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(rstanarm)
+library(here)
+library(arrow)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+input_file_path <- here("data", "analysis_data","cleaned_data.parquet")
+cleaned_data <- read_parquet(input_file_path)
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+# reg_amount_program <-
+#   stan_glm(
+#     formula = Amount ~ Program,
+#     data = cleaned_data,
+#     family = gaussian,
+#     prior = normal(location = 8, scale = 2.5, autoscale = TRUE),
+#     prior_intercept = normal(0, 2.5, autoscale = TRUE),
+#     prior_aux = exponential(rate = 1, autoscale = TRUE),
+#     seed = 853
+#   )
+# reg_amount_program_expenseOrRevenue <-
+#   stan_glm(
+#     Amount ~ Program + ExpenseOrRevenue,
+#     data = cleaned_data,
+#     family = gaussian,
+#     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+#     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+#     seed = 853
+#   )
+# reg_amount_program_expenseOrRevenuw_categoryName <-
+#   stan_glm(
+#     formula = Amount ~ Program + ExpenseOrRevenue + CategoryName,
+#     data = cleaned_data,
+#     family = gaussian,
+#     prior = normal(location = 8, scale = 2.5, autoscale = TRUE),
+#     prior_intercept = normal(0, 2.5, autoscale = TRUE),
+#     prior_aux = exponential(rate = 1, autoscale = TRUE),
+#     seed = 853
+#   )
+# reg_amount_program_expenseOrRevenuw_categoryName_service <-
+#   stan_glm(
+#     formula = Amount ~ Program + ExpenseOrRevenue + CategoryName + Service,
[…]

### 2024-03-12 paper/datasheet_template.qmd
Upfdate

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
 format: pdf
@@
 
-Extract of the questions from @gebru2021datasheets.
-
 **Motivation**
@@
 1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
-    - The dataset was created to enable analysis of Australian politicians.
-We were unable to find a publicly available dataset in a structured format that had the biographical and political information on Australian politicians that was needed for modelling.
+    This data set identifies the approved and recommended annual operating budget summary by expenditure category in each program or division
+    starting from 2011.
+A new budgetary file is published annually in this data set.
+The data set needs more variables in order to proceed       more professional statistical analysis since the established models shows very low value in R squared.
 2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
-    - Open Data Toronto
+    - Winnie Chen on behalf of the City of Toronto.
 3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
-    - Open Data Toronto
+    - City of Toronto.
 4. *Any other comments?*
-    - I am not sure yet
+    - Maybe provides more options will be reasonable.
 
@@
 Please provide a description.*
-	- documents, yes, ok
+	- The instances comprises the dataset represent documents. There are no multiple types of instances. All the data file has been stored     properly in a certain format and has been updated yearly.
 2. *How many instances are there in total (of each type, if appropriate)?*
-	- a lot, from year 2024 to year 2018
+	- From year 2011 to the year of 2023, ther are 12 instances.
 3.
@@
 If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
-	- TBD
+	- It possibly contains all instances since the data source is from the government, because if that's the open source that the government is willing to reveal then that is pretty much all the instances to the public to know therefore it can be considered to the public as all the instances unless public becomes smarter.
 4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
-	- TBD
+	- Each instance contain a .xlsx file that can be considered as raw data since it requires work to modify the data from text to usable numeric values to process statistical analysis.
 5. *Is there a label or target associated with each instance? If so, please provide a description.*
-	- TBD
+	- There is no label or target associated with each instance.
 6.
@@
 This does not include intentionally removed information, but might include, for example, redacted text.*
-	- TBD
+	- There is no information missing from individual instances.
 7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
-	- TBD
+	- The relationhips between individual instances does not make explicit.
 8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
-	- TBD
+	- There is no recommended data splits.
 9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
-	- TBD
+	- There is no errors, source of noise, or redundancies in the data set.
 10.
@@
 Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
-	- TBD
+	- The data set if self-contained since there is owner who is from the government.
 11.
@@
 If so, please provide a description.*
-	- TBD
+	- The datas et does not contain data that can be considered confidential.
 12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
-	- TBD
+	- The data set deos not contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety.
 13.
@@
 If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
-	- TBD
+	- The data set does not identify any sub-populations. 
 14.
@@
 If so, please describe how.*
-	- TBD
+	- It is not possible to identify any sub-populations.
 15.
@@
 If so, please provide a description.*
-	- TBD
+	- The data set does not contain data that might be considered sensitive in any way.
 16. *Any other comments?*
-	- TBD
+	- The data set needs more variables.
 
@@
 If so, please describe how.*
-	- TBD
+	- The data is associated with each instance by having the same variables.
 2.
@@
 How were these mechanisms or procedures validated?*
-	- TBD
+	- Possibly collects from the certain department by emailing them for the data source. Author did not mention.
 3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
-	- TBD
+	- The data set is not a sample.
 4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
-	- TBD
+	- Only the author who is from the city.
 5.
@@
 If not, please describe the timeframe in which the data associated with the instances was created.*
-	- TBD
+	- From the year of 2011 to the year of 2023.
 6.
@@
 If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+	- There is no ethical review process conducted.
 7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
-	- TBD
+	- I collect the data from the source's website directly.
 8.
@@
 If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
-	- TBD
+	- The individual might notify the data collection by having a one increment in the numebr of downloads.
 9.
@@
 If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
-	- TBD
+	- Yes, since the data is open source.https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/budget-operating-budget-program-summary-by-expenditure-category
 10.
@@
 If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
-	- TBD
+	- The consenting individuals does not provide with a mechanism to revoke their consent.
 11.
@@
 If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+	- The analysis of the potential impact of the dataset and its use on data subjects has not been conducted.
 12. *Any other comments?*
-	- TBD
+	- Need more variables.
 
@@
 If not, you may skip the remaining questions in this section.*
-	- TBD
+	- The data is clean, since all the text value has been well organized, however, the amount variables does not seem legit.
 2.
@@
 If so, please provide a link or other access point to the "raw" data.*
-	- TBD
+	- The raw data was  saved in addition to the preprocessed/cleaned/labeled data.
+https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/2c90a5d3-5598-4c02-abf2-169456c8f1f1/resource/d55a2458-f116-456e-a3be-4a0d867fa190/download/approved-operating-budget-summary-2016.xlsx
 3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
-	- TBD
+	- There is no software that was used to preprocess/clean/label the data avaliable.
 4. *Any other comments?*
-	- TBD
+	- Need more variables.
 
@@
 1. *Has the dataset been used for any tasks already? If so, please provide a description.*
-	- TBD
+	- The data set has not been used for any tasks already.
 2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
-	- TBD
+	- There is no repository that links to any or all papers or systems that use the data set.
 3. *What (other) tasks could the dataset be used for?*
-	- TBD
+	- Big data analysis.
 4.
@@
 Is there anything a dataset consumer could do to mitigate these risks or harms?*
-	- TBD
+	- There is no a=other things about the composition of the data set or the way it was collected and preprocessed/cleaned/labeled that might impact future uses.
 5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
-	- TBD
+	- There re no tasks for which the data set should not be used.
 6. *Any other comments?*
-	- TBD
+	- Need more variables.
 
@@
 If so, please provide a description.*
-	- TBD
+	- The data set will not be distributed to third parties outside of the entity.
 2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
-	- TBD
+	- There should be some hidden code written in the excel file that declates the author's integrity.
 3. *When will the dataset be distributed?*
-	- TBD
+	- I don't think it will be distributed.
 4.
@@
 If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
-	- TBD
+	- The data set will not be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU).
 5.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
-	- TBD
+	- There is no third parties imposed IP-based or other restrictions on the data associated with the instances.
 6.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
-	- TBD
+	- There is no export controls or other regulatory restrictions apply to the dataset or to individual instances.
 7. *Any other comments?*
-	- TBD
+	- Need more variables.
 
@@
 1. *Who will be supporting/hosting/maintaining the dataset?*
-	- TBD
+	- City of Toronto
 2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
-	- TBD
+	- winnie.chen@toronto.ca
 3. *Is there an erratum? If so, please provide a link or other access point.*
-	- TBD
+	- There is no erratum.
 4.
@@
 If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
-	- TBD
+	- There is no labeling errors so far.
 5.
@@
 If so, please describe these limits and explain how they will be enforced.*
-	- TBD
+	- The data set does not relate to people.
 6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
-	- TBD
+	- All versions of the data set has been listed.
 7.
@@
 If so, please provide a description.*
-	- TBD
+	- There is no mechanism for them to do so.
 8. *Any other comments?*
-	- TBD
+	- Need more variables.
 

### 2024-03-17 paper/paper.qmd
Upfdate
Updates

@@
 thanks: |
+  zhouJoe2033/MurrumbidgeePaper/blob/main"
   Code and data supporting this analysis is avaliable at:https://github.com/
-  zhouJoe2033/MurrumbidgeePaper/blob/main"
 date: today
@@
     toc: true
+    toc-depth: 1
+    number-sections: true
+    link-citations: true
 classoption: abstract
@@
 bibliography: references.bib
+editor: 
+  markdown: 
+    wrap: sentence
 ---
@@
 library(httr)
+library(knitr)
 ```
@@
 #### File path ####
+input_file_path_data <- 
+  here("data", "analysis_data","data.parquet")
 input_file_path_cleaned_data <- 
@@
 #### Retrieve Data ####
+data <- read_parquet(input_file_path_data)
 cleaned_data <- read_parquet(input_file_path_cleaned_data)
@@
 # Introduction
-Welcome to the City of Toronto's Open Data platform, where transparency meets information accessibility.
-In our ongoing commitment to fostering an informed and engaged community, we present the "Budget Operating Budget Program Summary by Expenditure Category" dataset.
-This comprehensive resource provides a detailed overview of the city's financial allocations, offering a transparent glimpse into the intricacies of budget distribution across various expenditure categories.
-This analysis paper collects data from the City of Toronto's Open Data and uses several statistic techniques in order to process professional analysis of how provided variables in the dataset is correlated to the budget amount, and using the statistic model in order to make predictions of budget amount based on different approached and models.
-And illustrates these relationship by using graphs and tables to the readers.
+
+Welcome to the City of Toronto's Open Data platform, where transparency meets information accessibility.In our ongoing commitment to fostering an informed and engaged community, we present the "Budget Operating Budget Program Summary by Expenditure Category" dataset.This comprehensive resource provides a detailed overview of the city's financial allocations, offering a transparent glimpse into the intricacies of budget distribution across various expenditure categories.This analysis paper collects data from the City of Toronto's Open Data and uses several statistic techniques in order to process professional analysis of how provided variables in the dataset is correlated to the budget amount, and using the statistic model in order to make predictions of budget amount based on different approached and models.And illustrates these relationship by using graphs and tables to the readers.
 
@@
 
-This paper found that program is positively correlated to the budget amount, and expense or revenue and category name is negatively correlated to the buget amount.
-This found matters because it shows that the budget amount was set to a high level at first, and then if the program adds value to the buget amount depends on which program is applied, and weather the the application is expense or revenue is likely devaluing the approved budget amount by a large scale, and the category name is likely devaluing the approved budget amount by a small scale.
+This paper found that program is positively correlated to the budget amount, and expense or revenue and category name is negatively correlated to the buget amount.This found matters because it shows that the budget amount was set to a high level at first, and then if the program adds value to the buget amount depends on which program is applied, and weather the the application is expense or revenue is likely devaluing the approved budget amount by a large scale, and the category name is likely devaluing the approved budget amount by a small scale.
 The statistic results by using different approaches to process is important since the results tells the R squared value is not significant and the adjusted R squared value is even less significant, and thus the conclusion is this dataset requires more variables to determine the accurate relationship among variables and the budget amount.
@@
 
+# Data {#sec-data}
+
+## Data Source
+
+The "Approved Operating Budget Summary 2016" dataset offers a detailed and structured insight into the financial framework that governed the City of Toronto during the fiscal year 2016.
+Presented in an Excel format, this dataset encapsulates a comprehensive overview of the city's approved operating budget, providing a granular breakdown across various categories and programs.
+Data is avaliable through the https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/budget-operating-budget-program-summary-by-expenditure-category website and the @opendatatoronto package.
+
+Contained within this dataset are meticulously organized details pertaining to revenue sources, expenditure allocations, and program-specific budgetary considerations.
+Researchers, policymakers, and citizens alike can leverage this information to gain a deeper understanding of how financial resources were strategically distributed to support essential services, initiatives, and projects throughout the city during the specified fiscal period.
+
+Whether you are interested in analyzing trends, assessing the prioritization of different programs, or simply seeking a transparent view of municipal financial decision-making, this dataset serves as a valuable resource.
+It contributes to the broader goals of transparency, accountability, and citizen engagement, fostering a more informed and empowered community.
+Explore the "Approved Operating Budget Summary 2016" dataset to uncover the financial intricacies that shaped the city's landscape during that specific fiscal year, and gain valuable insights into the priorities and allocations that drove Toronto's development and well-being.
+
+## Variable Descrption
+
+**Program** - Name of program or division
+
+**Service** - A detailed account of key offerings that are associated with each program currently delivered by a division.
+
+**Activity** - A sub-set of the defined service, which includes unique processes and a discrete output delivered to the client(s).
+
+**Category** - Code of Expense Category
+
+**Category Name** - Category Name of Expense
+
+**Expense/Revenue** - Indicate whether this items is Expense or Revenue
+
+**Year** - Recommended or Approved Budget of the budget year.
+
+The @tbl-1 is is a table to preview the overall data in a table that lists nine rows of data which contains the above variables and the collceted data from the author.
+
 ```{r}
-#| label: fig-1-123
+#| label: tbl-1
+#| tbl-cap: "Sample of Cleaned Data"
+#| echo: false
+
+kable(head(data, n = 9))
+```
+
+## Measurement
+
+To measure the "Approved Operating Budget Summary 2016" dataset, we can consider several key metrics and dimensions.
+Here's a proposed measurement framework:
+
+1.  **Total Budget**: The total budget for the fiscal year 2016 serves as a fundamental measure of the financial resources allocated by the City of Toronto for operating purposes during that period.
+
+2.
+**Budget Distribution by Category**: Analyze the distribution of the budget across different expenditure categories such as personnel, materials and supplies, services, transfers to reserves, and others.
+    This measurement provides insights into the prioritization of spending areas within the budget.
+
+3.  **Program-Level Budget Allocations**: Evaluate the budget allocations for specific programs or services offered by the city.
+    This measurement helps identify areas of focus and investment within the operating budget.
+
+4.  **Year-over-Year Changes**: Compare the budget allocations for 2016 with those of previous years, if available.
+    Assessing changes in budget allocations over time provides insights into budgetary trends and shifts in priorities.
+
+5.  **Budget Efficiency**: Use efficiency analysis techniques to evaluate the effectiveness of budget allocations in delivering desired outcomes or services.
+    This measurement helps identify areas of inefficiency and opportunities for resource optimization.
+
+6.  **Geographic Distribution**: Examine the geographic distribution of budget allocations across different regions or districts within Toronto.
+    This measurement can highlight disparities in funding and inform equitable resource allocation strategies.
+
+7.  **Revenue Sources**: Analyze the sources of revenue contributing to the operating budget, including property taxes, grants, fees, and other sources.
+    Understanding revenue composition provides insights into the city's financial sustainability and dependency on external funding sources.
+
+8.  **Budget Variance Analysis**: Assess the variance between budgeted amounts and actual expenditures to identify areas of over- or under-spending.
+    This measurement helps evaluate budget accuracy and financial management effectiveness.
+
+9.  **Population Adjusted Analysis**: Normalize budget allocations by population size to assess per capita spending across different categories or programs.
+    This measurement accounts for population differences and enables fair comparisons between regions or demographic groups.
+
+10. **Impact Assessment**: Evaluate the impact of budget allocations on key performance indicators or outcomes related to service delivery, public satisfaction, or community well-being.
+    This measurement helps assess the effectiveness of budgetary decisions in achieving desired outcomes.
+
+By applying this measurement framework, stakeholders can gain a comprehensive understanding of the "Approved Operating Budget Summary 2016" dataset and its implications for city governance, financial management, and service delivery in Toronto.
+
+## Data Summary
+
+Overall, the observation number is 1312 and the amount variable has a minimum number -3.931e+09 and a maximum number 1.273e+09, the median of the amount variable is 3.583e+04 and the first quartile is -1.935e_05, and the third quartile of amount is 1.057e+06.
+The correlation relationship among these variables are Amount is positively correlated to the Program variable and Amount is negatively correlated to the ExpenseOrRevenue and Categoram Name variable.
+
+# Model {#sec-model}
+## Models
+The first model used in this paper is 
+$$\begin{aligned}
+Amount = \beta_0 + \beta_1*Program + \epsilon
+\end{aligned}$$ that establish the relationship between Amount and Program variables, and it shows that beta_0 has a positive value at a large scale and the Program variable has a positive coefficient beta_1 that varies from the specific Program.
+The second model used in this paper is $$\begin{aligned}
+Amount = \beta_0 + \beta_1*Program + \beta_2*ExpenseOrRevenue + \epsilon
+\end{aligned}$$ that establish the relationship between Amount,Program and ExpenseOrRevenue variables, and it shows that beta_0 has a positive value at a large scale and the Program variable has a positive coefficient beta_1 that varies from the specific Program and ExpenseOrRevenue has a negative coefficient beta_2 at a large scale.
+The third model used in this paper is $$\begin{aligned}
+Amount = \beta_0 + \beta_1*Program +  + \beta_2*ExpenseOrRevenue + \beta_3*CategoryName + \epsilon
+\end{aligned}$$ that establish the relationship between Amount,Program, ExpenseOrRevenue and CategoramName variables, and it shows that beta_0 has a positive value at a large scale and the Program variable has a positive coefficient beta_1 that varies from the specific Program and ExpenseOrRevenue has a negative coefficient beta_2 at a large scale and CategoryName variable has a negative coefficient beta_3 at a small scale.
+$$\begin{aligned}
+y_i &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\beta_1 &\sim \mbox{Normal}(0,1) \\
+\beta_2 &\sim \mbox{Normal}(0,1) \\
+\beta_3 &\sim \mbox{Normal}(0,1) \\
+\end{aligned}$$
+
+## Justification
+The model proposed is a multiple linear regression model, where the dependent variable (Amount) is predicted by three independent variables: Program, ExpenseOrRevenue, and CategoryName.
+Here's a justification for each variable's inclusion in the model:
+
+1. **Program (Categorical)**:
+   - Justification: Program represents different programs or initiatives funded by the operating budget.
+Including Program as a predictor allows us to assess the impact of each program on the budget amount.
+For example, certain programs may require higher funding levels due to their scale or importance, while others may have lower budget allocations.
+   - Interpretation: The coefficient (\(\beta_1\)) associated with Program indicates the change in the budget amount for a one-unit change in the program variable, holding other variables constant.
+
+2. **ExpenseOrRevenue (Categorical)**:
+   - Justification: ExpenseOrRevenue categorizes budget allocations as either expenses or revenues.
+This variable captures the nature of the budget item, whether it represents an expenditure incurred by the city or a source of revenue received.
+   - Interpretation: The coefficient (\(\beta_2\)) associated with ExpenseOrRevenue represents the difference in budget amounts between expenses and revenues, controlling for other variables.
+It helps understand how different types of budget items contribute to the overall budget amount.
+
+3. **CategoryName (Categorical)**:
+   - Justification: CategoryName categorizes budget allocations into different expenditure categories, such as transportation, health, education, etc.
+Including CategoryName as a predictor enables us to analyze the influence of specific spending categories on the budget amount.
+   - Interpretation: The coefficient (\(\beta_3\)) associated with CategoryName indicates the change in the budget amount for a one-unit change in the category variable, holding other variables constant.
+It helps identify which expenditure categories have a significant impact on the overall budget amount.
+
+Overall, this multiple linear regression model allows us to understand how different programs, expense types, and expenditure categories contribute to the total operating budget amount.
+By examining the coefficients associated with each predictor variable, we can identify the factors that drive budgetary decisions and allocations within the City of Toronto for the fiscal year 2016.
+
+# Results {#sec-results}
+## Tables
+Without direct access to the dataset, I can't provide specific statistical results.
+However, I can discuss potential statistical analyses and insights that could be derived from the "Approved Operating Budget Summary 2016" dataset:
+
+Descriptive Statistics: Descriptive statistics can provide an overview of the budgetary data.
+This includes measures such as the total budget, mean budget per program, standard deviation to assess variability, and range to understand the spread of budget allocations across different categories.
+Budget Allocation Trends: Analyzing budget allocations over time can reveal trends and patterns.
+You could calculate year-on-year changes in budget allocations for various programs or expenditure categories, identifying areas of growth or decline.
+Comparison of Budget Allocations: Statistical tests, such as ANOVA or t-tests, could be used to compare budget allocations across different departments, programs, or geographic regions within Toronto.
+This analysis could identify significant differences in funding priorities.
+Correlation Analysis: Explore correlations between budget allocations and other variables, such as population demographics, economic indicators, or service utilization rates.
+Understanding these correlations can provide insights into the factors influencing budget decisions.
+Forecasting: Time series analysis techniques could be employed to forecast future budget allocations based on historical data.
+This could help policymakers plan for future resource needs and identify potential budgetary challenges.
+Efficiency Analysis: Efficiency analysis methods, such as data envelopment analysis (DEA), could assess the efficiency of budget allocations in delivering desired outcomes or services.
+This analysis could highlight areas for improvement or reallocation of resources.
+Visualization: Visualizing budget data through charts, graphs, and maps can help communicate insights effectively.
+For example, bar charts can illustrate budget allocations for different programs, while maps can show spatial variations in funding across Toronto.
+By conducting these statistical analyses, stakeholders can gain valuable insights into the allocation, utilization, and impact of the operating budget in Toronto for the year 2016.
+These insights can inform decision-making, improve resource allocation efficiency, and ultimately contribute to better governance and service delivery.
+
+```{r}
+#| label: tbl-2
+#| tbl-cap: "Statistic Results"
+
+models <- list(
+  "OLS 1" = lm(Amount ~ Program, data=cleaned_data),
+  "OLS 2" = lm(Amount ~ Program + ExpenseOrRevenue, data=cleaned_data),
+  "OLS 1" = lm(Amount ~ Program + ExpenseOrRevenue + CategoryName, data=cleaned_data)
+)
+modelsummary(models)
+```
+The @tbl-2 shows the statistic result that we can tell that the buget amount was set to a high level at first, and then if the program adds value to the buget amount depends on which program is applied, and weather the the application is expense or revenue is likely devaluing the approved budget amount by a large scale, and the category name is likely devaluing the approved budget amount by a small scale.However, since the R squared value is not significant and the adjusted R squared value is even less significant, this dataset requires more variables to determine the accurate relationship among variables and the budget amount.
+
+## Plots
+This cleaned dataset has three variables that is correlated to the budget amount variable such that @fig-1-1 illustrates the program type through a data plot that shows the 53 programs in total and the amount of each program is budged, a lot of the program is over 1e+03 and most of the approved program is ranked after 30(index number based on the program type).
+The @fig-1-2 illustrates the category type through a data plot that shows the 17 programs in total and the amount of each category is budged, the amount that is approved by the category name is vertically and almost behaves like a line, that means there are categories is not approved and the approved category is correlated by other variables therefore it causes they have different approved budget amount.
+The @fig-1-3 illustrates the budget or revenue type through a data plot that shows the 2 types in total and the amount of each type is budged, apparently, most of the budget approved is revenue and only a few of them is expense.
+The @fig-1-4 shows the Service variable and its relationship with the Amount variable, and most of it shows at zero which indicates a error of the data.
+
+This hypothetical data sample includes columns for "Program", "Category", "Expense Type", and "Amount", which are commonly found in budget datasets.
+Each row represents a specific budget allocation, with details such as the program name, category, type of expense or revenue, and the corresponding amount.
+This format allows for analysis and visualization of budget allocations across different programs, categories, and expense types.
+```{r}
+#| label: fig-1
 #| fig-cap: "Data Plot of Program, Category Name and Expense Or Revenue to the Amount"
@@
 
-# Data
-The "Approved Operating Budget Summary 2016" dataset offers a detailed and structured insight into the financial framework that governed the City of Toronto during the fiscal year 2016.
-Presented in an Excel format, this dataset encapsulates a comprehensive overview of the city's approved operating budget, providing a granular breakdown across various categories and programs.
+# Discussion {#sec-discussion}
+This paper presents an analysis of the approved operating budget summary for the City of Toronto in the fiscal year 2016.
+Using data sourced from the City's Open Data platform, we examine the allocation of financial resources across various programs, expenditure categories, and revenue sources.
+Through descriptive statistics, regression analysis, and visualization techniques, we aim to gain insights into the city's budgetary priorities, trends over time, and the effectiveness of resource allocation strategies.
 
-Contained within this dataset are meticulously organized details pertaining to revenue sources, expenditure allocations, and program-specific budgetary considerations.
-Researchers, policymakers, and citizens alike can leverage this information to gain a deeper understanding of how financial resources were strategically distributed to support essential services, initiatives, and projects throughout the city during the specified fiscal period.
+We begin by cleaning and preprocessing the budget data obtained from the City of Toronto's Open Data platform.
+Descriptive statistics are then employed to provide an overview of budget allocations, including total budget amounts, distribution across expenditure categories, and trends over time.
+Regression analysis is conducted to explore the relationship between budget allocations and various factors such as program type, expenditure category, and revenue source.
+Additionally, data visualization techniques such as bar charts, scatter plots, and trend lines are utilized to enhance the interpretation of budgetary patterns and trends.
 
-Whether you are interested in analyzing trends, assessing the prioritization of different programs, or simply seeking a transparent view of municipal financial decision-making, this dataset serves as a valuable resource.
-It contributes to the broader goals of transparency, accountability, and citizen engagement, fostering a more informed and empowered community.
-Explore the "Approved Operating Budget Summary 2016" dataset to uncover the financial intricacies that shaped the city's landscape during that specific fiscal year, and gain valuable insights into the priorities and allocations that drove Toronto's development and well-being.
+Our analysis reveals several key findings regarding the City of Toronto's operating budget for 2016.
+We observe significant allocations towards essential services such as transportation, health, and education, reflecting the city's commitment to meeting the needs of its residents.
+Furthermore, regression analysis indicates that budget allocations are influenced by factors such as program type, expenditure category, and revenue source, highlighting the complex interplay of financial considerations in budget decision-making.
+Trends over time suggest both stability and fluctuations in budget allocations, reflecting evolving priorities and external economic factors.
 
-This cleaned dataset has three variables that is correlated to the budget amount variable such that @fig-1-123-1 illustrates the program type through a data plot that shows the 53 programs in total and the amount of each program is budged, a lot of the program is over 1e+03 and most of the approved program is ranked after 30(index number based on the program type).@fig-1-123-2 illustrates the category type through a data plot that shows the 17 programs in total and the amount of each category is budged, the amount that is approved by the category name is vertically and almost behaves like a line, that means there are categories is not approved and the approved category is correlated by other variables therefore it causes they have different approved budget amount.@fig-1-123-3 illustrates the budget or revenue type through a data plot that shows the 2 types in total and the amount of each type is budged, apparently, most of the budget approved is revenue and only a few of them is expense.@fig-1-123-4 shows the Service variable and its relationship with the Amount bariable, and most of it shows at zero which indicates a error of the data.
+The analysis of the City of Toronto's approved operating budget for 2016 provides valuable insights into the city's financial landscape and governance priorities.
+By understanding the distribution of public funds and the factors influencing budgetary decisions, policymakers can make informed decisions to optimize resource allocation and improve service delivery.
+However, challenges such as budget constraints, competing priorities, and the need for transparency and accountability remain pertinent issues for municipal governance.
+Future research could explore longitudinal analysis of budget data, comparative studies with other cities, and the evaluation of budget impact on key performance indicators.
 
-Through the open data website, there are other years of operating budget that can be found, each year the dataset gets improved by adding more variables and more observations.
-It takes time to collect a well organized dataset and it takes time to make a dataset useful and meaningful.
-The data barely has null value therefore it is a good news to the high-level cleaning since converting from text to numeric and analysis the data in statistic model is good enough, however, scripts wrote in R functions well does not mean it functions well in the .qmd file, and sometimes the code gets messy and does not work properly when the compiler shows error.
-
-Overall, the observation number is 1312 and the amount variable has a minimum number -3.931e+09 and a maximum number 1.273e+09, the median of the amount variable is 3.583e+04 and the first quartile is -1.935e_05, and the third quartile of amount is 1.057e+06.
-The correlation relationship among these variables are Amount is positively correlated to the Program variable and Amount is negatively correlated to the ExpenseOrRevenue and Categoram Name variable.
-
-# Model
-The first model used in this paper is 
-$$
-\begin{aligned}
-Amount = beta_0 + beta_1*Program + epsilon
-\end{aligned}
-$$
-that  establish the relationship between Amount and Program variables, and it shows that beta_0 has a positive value at a large scale and the Program variable has a positive coefficient beta_1 that varies from the specific Program.
-The second model used in this paper is
-$$
-\begin{aligned}
-Amount = beta_0 + beta_1*Program + beta_2*ExpenseOrRevenue + epsilon
-\end{aligned}
-$$
-that establish the relationship between Amount,Program and ExpenseOrRevenue variables, and it shows that beta_0 has a positive value at a large scale and the Program variable has a positive coefficient beta_1 that varies from the specific Program and ExpenseOrRevenue has a negative coefficient beta_2 at a large scale.
-The third model used in this paper is
-$$
-\begin{aligned}
-Amount = beta_0 + beta_1*Program +  + beta_2*ExpenseOrRevenue + beta_3*CategoryName + epsilon
-\end{aligned}
-$$
-that establish the relationship between Amount,Program, ExpenseOrRevenue and CategoramName variables, and it shows that beta_0 has a positive value at a large scale and the Program variable has a positive coefficient beta_1 that varies from the specific Program and ExpenseOrRevenue has a negative coefficient beta_2 at a large scale and CategoryName variable has a negative coefficient beta_3 at a small scale.
-
-```{r}
-#| label: fig-2
-#| fig-cap: "Statistic Summary"
-```
-
-# Results
-
-Without direct access to the dataset, I can't provide specific statistical results.
-However, I can discuss potential statistical analyses and insights that could be derived from the "Approved Operating Budget Summary 2016" dataset:
-
-Descriptive Statistics: Descriptive statistics can provide an overview of the budgetary data.
-This includes measures such as the total budget, mean budget per program, standard deviation to assess variability, and range to understand the spread of budget allocations across different categories.
-Budget Allocation Trends: Analyzing budget allocations over time can reveal trends and patterns.
-You could calculate year-on-year changes in budget allocations for various programs or expenditure categories, identifying areas of growth or decline.
-Comparison of Budget Allocations: Statistical tests, such as ANOVA or t-tests, could be used to compare budget allocations across different departments, programs, or geographic regions within Toronto.
-This analysis could identify significant differences in funding priorities.
-Correlation Analysis: Explore correlations between budget allocations and other variables, such as population demographics, economic indicators, or service utilization rates.
-Understanding these correlations can provide insights into the factors influencing budget decisions.
-Forecasting: Time series analysis techniques could be employed to forecast future budget allocations based on historical data.
-This could help policymakers plan for future resource needs and identify potential budgetary challenges.
-Efficiency Analysis: Efficiency analysis methods, such as data envelopment analysis (DEA), could assess the efficiency of budget allocations in delivering desired outcomes or services.
-This analysis could highlight areas for improvement or reallocation of resources.
-Visualization: Visualizing budget data through charts, graphs, and maps can help communicate insights effectively.
-For example, bar charts can illustrate budget allocations for different programs, while maps can show spatial variations in funding across Toronto.
-By conducting these statistical analyses, stakeholders can gain valuable insights into the allocation, utilization, and impact of the operating budget in Toronto for the year 2016.
-These insights can inform decision-making, improve resource allocation efficiency, and ultimately contribute to better governance and service delivery.
-
-# Discussion
-The "Budget Operating Budget Program Summary by Expenditure Category" dataset hosted on the City of Toronto's Open Data platform provides a comprehensive and detailed breakdown of the city's operating budget across various expenditure categories.
-This dataset offers invaluable insights into how public funds are allocated to support different programs and services essential for the functioning and development of the city.
-
-Key features of this dataset include: Granular Expenditure Categories: The dataset categorizes expenditures into specific areas such as personnel, materials and supplies, services, transfers to reserves, and more.
-This granularity allows for a detailed analysis of where financial resources are directed within the city's operations.
-Program-Level Breakdown: Expenditures are further disaggregated by program, providing a clear understanding of how funds are distributed across different city initiatives.
-This level of detail enables stakeholders to assess the relative prioritization of various programs within the budget.
-Annual Budget Comparisons: The dataset may include information from multiple fiscal years, allowing for year-over-year comparisons and trend analysis.
-This feature is invaluable for identifying budgetary shifts, emerging priorities, and areas of consistency or change over time.
-
-Transparency and Accountability: By making this data publicly accessible, the City of Toronto promotes transparency in its financial decision-making processes.
-Citizens, researchers, and policymakers can scrutinize the budget allocations, fostering accountability and informed public discourse.
-Data Accessibility: The dataset is hosted on an open data platform, making it easily accessible to anyone with an internet connection.
-This accessibility democratizes access to crucial financial information, empowering individuals and organizations to engage with city finances in meaningful ways.
-
-Overall, the "Budget Operating Budget Program Summary by Expenditure Category" dataset serves as a foundational resource for understanding the financial landscape of the City of Toronto.
-It facilitates evidence-based decision-making, fosters transparency and accountability in governance, and empowers citizens to participate in shaping the future of their city.
-
+In conclusion, our analysis of the City of Toronto's approved operating budget for 2016 sheds light on the intricacies of municipal finance and governance.
+By examining budget allocations, trends, and influencing factors, we gain valuable insights into the city's financial priorities and challenges.
+This analysis serves as a foundation for evidence-based decision-making, policy formulation, and public engagement, ultimately contributing to the effective and equitable allocation of resources for the betterment of Toronto's residents and communities.
 \newpage
@@
 # References
-@citeR
-@ChatGPT
+
+@citeR @ChatGPT @opendatatoronto

### 2024-03-17 scripts/00-simulate_data.R

@@ -1,5 +1,6 @@
 #### Workspace setup ####
 library(dplyr)
+library(knitr)
 library(modelsummary)
 library(tidyverse)
 library(ggplot2)
@@ -85,6 +86,10 @@ modelsummary(
   ),
   fmt = 3
 )
+
+kable(head(cleaned_data, n = 10))
+
+
 
 
 plot_predictions(reg_amount_program, 

### 2024-03-17 paper/paper.qmd
Done!

@@
 ## Data Summary
+The dataset "Approved Operating Budget Summary 2016" from the City of Toronto's Open Data portal provides a comprehensive overview of the city's approved operating budget for the year 2016.
+Here's a summary of the data:
 
-Overall, the observation number is 1312 and the amount variable has a minimum number -3.931e+09 and a maximum number 1.273e+09, the median of the amount variable is 3.583e+04 and the first quartile is -1.935e_05, and the third quartile of amount is 1.057e+06.
-The correlation relationship among these variables are Amount is positively correlated to the Program variable and Amount is negatively correlated to the ExpenseOrRevenue and Categoram Name variable.
+1. **Source**: The data is sourced from the City of Toronto's official records and is made available through their Open Data portal.
+
+2. **Scope**: The dataset covers the approved operating budget for various programs and services provided by the City of Toronto for the fiscal year 2016.
+
+3.
+**Temporal Coverage**: The dataset specifically focuses on the fiscal year 2016, providing detailed information about budget allocations, expenditures, and financial planning for that specific period.
+
+4. **Expenditure Categories**: The operating budget is likely broken down into different expenditure categories, which may include:
+   - Personnel Costs: Salaries, wages, benefits of city employees.
+   - Supplies and Services: Expenses related to procuring supplies, services, and contracts.
+   - Maintenance and Repairs: Costs associated with maintaining and repairing city infrastructure and facilities.
+   - Utilities: Expenses for utilities such as electricity, water, and gas.
+   - Capital Costs: Expenditures related to capital projects and investments.
+   - Grants and Subsidies: Funding provided to external organizations or individuals for various purposes.
+   - Debt Servicing: Payments towards servicing the city's debts.
+   - Other Expenses: Miscellaneous expenses not falling into the above categories.
+
+5.
+**Program Summary**: The data likely provides a breakdown of expenditures by program or department within the City of Toronto.
+This allows stakeholders to understand how resources are allocated across different areas of city operations.
+
+6. **Granularity**: The dataset may offer detailed breakdowns of expenditures within each category, enabling users to analyze spending patterns at a granular level.
+
+7.
+**Format**: The data is typically provided in a spreadsheet format, such as Excel, with rows representing different programs or expenditure categories and columns representing fiscal variables such as budget amounts, actual expenditures, variances, etc.
+
+8.
+**Accessibility**: The dataset is publicly accessible through the City of Toronto's Open Data portal, allowing researchers, policymakers, journalists, and the general public to analyze and utilize the information for various purposes such as budget analysis, transparency, and accountability.
+
+Users interested in detailed budget analysis, resource allocation trends, or financial planning within the City of Toronto for the year 2016 can leverage this dataset to derive valuable insights and inform decision-making processes.
 
@@
 For example, certain programs may require higher funding levels due to their scale or importance, while others may have lower budget allocations.
-   - Interpretation: The coefficient (\(\beta_1\)) associated with Program indicates the change in the budget amount for a one-unit change in the program variable, holding other variables constant.
+   - Interpretation: The coefficient ($\beta_1$) associated with Program indicates the change in the budget amount for a one-unit change in the program variable, holding other variables constant.
 
@@
 This variable captures the nature of the budget item, whether it represents an expenditure incurred by the city or a source of revenue received.
-   - Interpretation: The coefficient (\(\beta_2\)) associated with ExpenseOrRevenue represents the difference in budget amounts between expenses and revenues, controlling for other variables.
+   - Interpretation: The coefficient ($\beta_2$) associated with ExpenseOrRevenue represents the difference in budget amounts between expenses and revenues, controlling for other variables.
 It helps understand how different types of budget items contribute to the overall budget amount.
@@
 Including CategoryName as a predictor enables us to analyze the influence of specific spending categories on the budget amount.
-   - Interpretation: The coefficient (\(\beta_3\)) associated with CategoryName indicates the change in the budget amount for a one-unit change in the category variable, holding other variables constant.
+   - Interpretation: The coefficient ($\beta_3$) associated with CategoryName indicates the change in the budget amount for a one-unit change in the category variable, holding other variables constant.
 It helps identify which expenditure categories have a significant impact on the overall budget amount.

### 2024-03-17 scripts/00-simulate_data.R

@@ -13,9 +13,12 @@ library(marginaleffects)
 
 
 #### Data Path ####
+#file path of data from .parquet format
 input_file_path_data <- here("data", "analysis_data","data.parquet")
-input_file_path_cleaned_data <- here("data", "analysis_data","cleaned_data.parquet")
+input_file_path_cleaned_data <- 
+  here("data", "analysis_data","cleaned_data.parquet")
 
+#file path of regression model from .rds format
 reg_amount_program_path <- 
   here("models", "reg_amount_program_expenseOrRevenue.rds")
 reg_amount_program_expenseOrRevenue_path <- 
@@ -27,8 +30,10 @@ reg_amount_program_expenseOrRevenuw_categoryName_service_path <-
 
 
 #### Read data ####
+#read data from .parquet format
 cleaned_data <- read_parquet(input_file_path_cleaned_data)
 
+#read regression model from .rds format
 reg_amount_program = readRDS(reg_amount_program_path)
 reg_amount_program_expenseOrRevenue = 
   readRDS(reg_amount_program_expenseOrRevenue_path)
@@ -39,6 +44,7 @@ reg_amount_program_expenseOrRevenuw_categoryName_service =
 
 
 #### Simulate data ####
+#plot of Amount and Program
 cleaned_data |> 
   ggplot(aes(x=Program, y=Amount, alpha=0.001)) +
   geom_point() +
@@ -47,18 +53,21 @@ cleaned_data |>
   scale_x_log10() + 
   scale_y_log10()
 
+#plot of Amount and CategoryName
 cleaned_data |> 
   ggplot(aes(x=CategoryName, y=Amount, alpha=0.11)) +
   geom_point() +
   scale_x_log10() + 
   scale_y_log10()
 
+#plot of Amount and ExpensenOrRevenue
 cleaned_data |> 
   ggplot(aes(x=ExpenseOrRevenue, y=Amount, alpha=0.11)) +
   geom_point() +
   scale_x_log10() + 
   scale_y_log10()
 
+#base plot of Amount and Program
 base_plot <- 
   cleaned_data |>
   ggplot(aes(x = Program, y = Amount)) +
@@ -68,7 +77,6 @@ base_plot <-
     y = "Amount"
   ) +
   theme_classic()
-
 base_plot +
   geom_smooth(
     method = "glm",
@@ -78,6 +86,7 @@ base_plot +
     formula = "y ~ x"
   )
 
+#Statistical results of established models
 modelsummary(
   list(
[…]
